Modeling and analyzing the Corona-virus warning app with the Isabelle Infrastructure framework

### 2020-07-10 paper.tex
started writing section on Refinement application

\section{Model for Corona-warning app}
\label{sec:model}

\subsection{Infrastructures, Policies, and Actors}
\label{sec:infra}
% Adapted from CSF paper
The Isabelle Infrastructure framework supports the representation of infrastructures 
as graphs with actors and policies attached to nodes. These infrastructures 
are the {\it states} of the Kripke structure. % for the attack trees.

The transition between states is triggered by non-parametrized
actions \texttt{get}, \texttt{move}, and \texttt{put} 
executed by actors. 
Actors are given by an abstract type \texttt{actor} and a function 
\texttt{Actor} that creates elements of that type from identities 
(of type \texttt{string} written \texttt{''s''} in Isabelle). 
Actors are contained in an infrastructure graph type \texttt{igraph}
constructed by its constructor \texttt{Lgraph}. 
\begin{ttbox}
{\bf datatype} igraph = 
         Lgraph (location \tttimes location)set 
                 location \ttfun identity set
                 identity \ttfun (string set \tttimes string set \tttimes efid)  
                 location \ttfun string \tttimes (dlm \tttimes data) set
                 location \ttfun efid set
                 actor \ttfun location \ttfun (identity \tttimes efid) set

\end{ttbox}
In the current application of the framework to the Corona-warning app case study, 
this graph contains a set of location pairs representing the %structure 
topology of the infrastructure
as a graph of nodes and a function%\footnote{ We use the common $\lambda$-abstraction, 
%e.g. \texttt{\ttlam x. True}, to define functions with parameters, here the 
%function returning True for any input $x$.} 
that assigns a set of actor identities to each node (location) in the graph.
The third component of an \texttt{igraph} assigns the credentials to each actor: 
a triple-valued function whose first range component is a set describing the credentials 
in the possession of an actor and the second component 
is a set defining the roles the actor can take on; 
most prominently the third component is the \texttt{efid} assigned to the actor. This is initially
just a natural number but will be refined to actually represent lists of ``Ephemeral'' Ids later
when refining the specification.
\begin{ttbox}
{\bf datatype} efid = Efid nat
\end{ttbox}
The fourth component of the type \texttt{igraph}
assigns security labelled data to locations, a feature not used in the
current application.
The second to last component assigns the set of efids of all currently present smart phones
to each location of the graph. The last component finally denotes the knowledge set of each
actor for each location: a set of pairs of actors and potential ids.

@@

In addition, as we can observe in the following example in the second refinement
step, we also have a simple implicit version of {\it action refinement}. In an
action refinement, traces may be refined by combining consecutive system events
into atomic events thereby reducing traces. This will be detailed on the example
below.

\subsection{Refining the Specification}
\label{sec:corref}

>>>>

\section{Modeling and analyzing Corona-warning app}
\label{sec:model}
\subsection{Infrastructures, Policies, and Actors}
\label{sec:infra}
% Adapted from CSF paper
The Isabelle Infrastructure framework supports the representation of infrastructures 
as graphs with actors and policies attached to nodes. These infrastructures 
are the {\it states} of the Kripke structure. % for the attack trees.

The transition between states is triggered by non-parametrized
actions \texttt{get}, \texttt{move}, and \texttt{put} 
executed by actors. 
Actors are given by an abstract type \texttt{actor} and a function 
\texttt{Actor} that creates elements of that type from identities 
(of type \texttt{string} written \texttt{''s''} in Isabelle). 
Actors are contained in an infrastructure graph type \texttt{igraph}
constructed by its constructor \texttt{Lgraph}. 
\begin{ttbox}
{\bf datatype} igraph = 
         Lgraph (location \tttimes location)set 
                 location \ttfun identity set
                 identity \ttfun (string set \tttimes string set \tttimes efid)  
                 location \ttfun string \tttimes (dlm \tttimes data) set
                 location \ttfun efid set
                 actor \ttfun location \ttfun (identity \tttimes efid) set
\end{ttbox}
In the current application of the framework to the Corona-warning app case study, 
this graph contains a set of location pairs representing the %structure 
topology of the infrastructure
as a graph of nodes and a function%\footnote{ We use the common $\lambda$-abstraction, 
%e.g. \texttt{\ttlam x. True}, to define functions with parameters, here the 
%function returning True for any input $x$.} 
that assigns a set of actor identities to each node (location) in the graph.
The third component of an \texttt{igraph} assigns the credentials to each actor: 
a triple-valued function whose first range component is a set describing the credentials 
in the possession of an actor and the second component 
is a set defining the roles the actor can take on; 
most prominently the third component is the \texttt{efid} assigned to the actor. This is initially
just a natural number but will be refined to actually represent lists of ``Ephemeral'' Ids later
when refining the specification.
\begin{ttbox}
{\bf datatype} efid = Efid nat
\end{ttbox}
The fourth component of the type \texttt{igraph}
assigns security labelled data to locations, a feature not used in the
current application.
The second to last component assigns the set of efids of all currently present smart phones
to each location of the graph. The last component finally denotes the knowledge set of each
actor for each location: a set of pairs of actors and potential ids.

@@

In addition, as we can observe in the following example in the second refinement
step, we also have a simple implicit version of {\it action refinement}. In an
action refinement, traces may be refined by combining consecutive system events
into atomic events thereby reducing traces. This will be detailed in the second refinement step
of the Corona-virus warning app considered next.

\subsection{Refining the Specification}
\label{sec:corref}
Clearly, fixed Ephemeral Ids are not really ephemeral. The model presented
in Section \ref{sec:model} is deliberately been chosen very abstractly to allow focusing on
the basic system architecture and finding an initial deanonymization attack.
We now introduce ``proper'' Ephemeral Ids and show how the system datatype can be refined to a system
that uses those instead of the fixed ones.

For the DP-3T Ephemeral Ids \TODOref{DP-3T Whitepaper}, for each day $t$ a seed $SK_t$
is used to generate a list of length $n = 24 * 60 / L$, where $L$ is the duration for which
the Ephemeral Ids are posted by the smart phone
\begin{ttbox}
  EphId1 || ... || EphIdn = PRG(PRF(SKt,``broadcast key''))
\end{ttbox} 
``where PRF is a pseudo-random function (e.g., HMAC-SHA256), ``broadcast key'' is a
fixed, public string, and PRG is a pseudorandom generator (e.g. AES in counter mode)'' \TODOref{DP-3T Whitepaper}.

From a cryptographic point of view, the crucial properties of the Ephemeral Ids are that
they are purely random, therefore, they cannot be guessed, but at the same time if -- after
the actual encounter between sender and receiver -- the seed $SK_t$ is published, it is feasible
to relate any of the \texttt{EphIDi} to $SK_t$  for all \texttt{i} $\in \{1, \ldots, n\}$.
For a formalisation of this crucial cryptographic property in Isabelle it suffices to define a new type
of list of Ephemeral Ids \texttt{efidlist} containing the root $SK_t$ (the first \texttt{efid}),
a current \texttt{efid} indicated by a list pointer of type \texttt{nat}, and the actual list of
\texttt{efid}s.
\begin{ttbox}
{\bf{datatype}} efidlist = Efids "efid" "nat" "efid list"
\end{ttbox}
We define functions for this datatype: \texttt{efidsroot} returning the first of the three
consituents in an \texttt{efidlist} (the root $SK_t$); \texttt{efids\_index} giving the second
component, the index of the current \texttt{efid}; \texttt{efids\_inc\_ind} applied to an
\texttt{efidlist} increments the index; \texttt{efids\_cur} returning the current \texttt{efid}
from the list and \texttt{efids\_list} for the entire list (the third component).
